---
title: "GMACS code notes"
---

```{r}
#| echo: false
`%+=%` = function (lhs, rhs) 
{
    lhs <- substitute(lhs)
    rhs <- substitute(rhs)
    kind <- 2L
    lazy <- TRUE
    env <- parent.frame()
    lhs+rhs;
}
```

## Model time frame-related variables
   * "terminal year": final complete model year (the OFL would be determined for terminal year + 1)
   * `syr`: model start year  [data file]
   * `nyr`: nominal model end year/terminal year (but not necessarily - depends on `nyrRetro` and `nproj`) [data file]
   * `nyrRetro1`: integer set by commandline "-retro x" (-1 if not set)  [commandline]
   * `nyrRetro`: integer set in gmacs.dat file to specify number of peels to drop [dat, not data, file]
     - overridden by `nyrRetro1`, if it is set
     - after being read from gmacs.dat (and possibly set to the value of `nyrRetro1`), it is saved as `nyrRetroNo`, then converted to $nyrRetro = nyr-nyrRetro$, the terminal year of the retrospective peel
   * `nyrRetroNo`: original value (after `nyrRetro1`check ) of `nyrRetro`

   * `pyr`: terminal year for projections  [prj file]
   * `nproj`: $nproj = pyr - nyr$
   * `Nyears`: $Nyears =(nyrRetro-syr+1)$
   * `Nyears2`: $Nyears2 =(nyrRetro+1-syr+1)$
   * `Nyears3`: $Nyears3 = 2*nfleet*nsex*(nyr+1-syr+1)$
   * `Nyears4`: $Nyears4 =(nyrRetro+nproj+1-syr+1)$
   * `nprojVec`: $nprojVec = max(2,nproj+1)$

   * `Env_vars_used`(syr,nyr+nproj,1,Num_env_vars+1)
   * `Rec_bias`(syr,nyr)
   * `model_controls`(1): first year of rec dev estimates
   * `rdv_syr`:  $rdv_syr = int(model_controls(1))$
   * `model_controls`(2): last year of rec dev estimates
   * `rdv_syr`: rdv_eyr = int(model_controls(2) -= (nyr-nyrRetro))$  **why subtracting here: clearer to subtract `nyrRetroNo`*

   * `CustomSelex`(1,Nyears3,1,nclass)
   * setting `CustomSelex`:
      - loop over `it`=1:2; `h`=1:nsex; `k`=1:nfleet; 
      - `offset` $= (it-1)*nfleet*nsex*(nyr+1-syr+1)+nfleet*(nyr+1-syr+1)*(h-1)+(nyr+1-syr+1)*(k-1)$
      - loop over `iyr`=syr:nyr+1; `ii`=1:nclass
        - *(ad_comm::global_datafile) >> `CustomSelex`(offset+iyr-syr+1,ii);

In the following, `NVarPar` is a counter for the number of std variables.

```{r, eval=FALSE}
  if (rdv_phz > 0) NVarPar %+=% (rdv_eyr-rdv_syr+1);      #<1>
  if (rec_prop_phz > 0) NVarPar %+=% (rdv_eyr-rdv_syr+1); #<2>
```

1. adds number of rec devs to counter for number of std variables
2. adds number of rec proportion devs to counter for number of std variables

### Calculated sdreport variables

Note that the following implied time dimensions (`syr`:`nyr`) are probably wrong if nyrRetroNo $\ne$ 0 and should be changed to imply `syr`:`nyrRetro` (i.e., `Nyears`).
```{r, eval=FALSE}
  NRefPars = NVarPar;                                #<1>
  if (OutRefPars==YES) NVarPar %+=% (6 + 3*nfleet);    #<2>
  NRecPar = NVarPar+1;                               #<3>
  if (OutRecruit==YES) NVarPar %+=% nsex*(nyr-syr+1);  #<4>
  NSSBPar = NVarPar+1;                               #<5>
  if (OutSSB==YES) NVarPar %+=% (nyr-syr+1);           #<6>
  NfbarPar = NVarPar+1;                              #<7>
  if (Outfbar==YES) NVarPar %+=% (nyr-syr+1);          #<8>
  NB0Par = NVarPar+1;                                #<9>
  if (OutDynB0==YES) NVarPar %+=% (nyr-syr+1);         #<10>
  cout << "Number of std variables: " << NVarPar << endl; #<11>
```

1. Saves pointer to potential start of management reference quantitites in std variables vector
2. adds management reference quantities to number of std variables
3. Saves pointer to potential start of sex-specific recruitment values in std variables vector
4. adds number of sex-specific recruitment values to number of std variables
5. Saves pointer to potential start of MMB values in std variables vector
6. adds number of MMB values to number of std variables
7. Saves pointer to potential start of `fbar`(?) values in std variables vector
8. adds number of `fbar`(?) values to number of std variables
9. Saves pointer to potential start of dynamic B0 values in std variables vector
10. adds number of dynamic B0 values to number of std variables
11. at end of section, `NVarPar` is the total number of std variables

```{r, eval=FALSE}
  vector logRbar(syr,nyrRetro+nproj+1); #<1>
  matrix histcat(1,2+nfleet,syr,nyr); #<2>
  `4darray` numbers_proj_gytl(1,n_grp,1,nprojVec,1,nseason,1,nclass); #<3>
  vector dyn_Bzero(syr,nyrRetro); #<4>
  for (int iyr=syr;iyr<=nyr;iyr++) M(h,m,iyr) = elem_prod(M(h2,m2,iyr),Mmult); #<5>
```

1. `logRbar` is dimensioned syr:nyrRetro+nproj+1
2. `histcat` is time-dimensioned syr:nyr
3. `numbers_proj_gytl` is the N matrix for the projection (w/ dims roup, year, season, length class); time-dimensioned 1:nprojVec
4. `dyn_Bzero` is dimensioned syr:nyrRetro here, but NVarPar treats it like its dimensioned syr:nyr (see above)

## Setting selectivity functions

The following code sets 
```{r, eval=FALSE}
  int nsel_pnt = 0;
  for (int it=1;it<=2;it++){
    for (int h=1;h<=nsex;h++){
      for (int k=1;k<=nfleet;k++){
        int ipnt2 = (it-1)*nsex+h;
        switch ( slx_type_in(ipnt2,k) ){
          case SELEX_UNIFORM1:                                  ///> uniform 1
            for (int iyr=syr;iyr<=nyrRetro+nprojUse+1;iyr++){ #<1>
              if (it==1) for (int i=1;i<=nclass;i++) { log_slx_capture(k,h,iyr,i) = log(1.0-TINY);}
              if (it==2) for (int i=1;i<=nclass;i++) { log_slx_retaind(k,h,iyr,i) = log(1.0-TINY); log_slx_discard(k,h,iyr,i) = log(TINY); }
            }
            break;
          case SELEX_UNIFORM0:                                  ///> uniform 0
            for (int iyr=syr;iyr<=nyrRetro+nprojUse+1;iyr++){ #<1>
              if (it==1) for (int i=1;i<=nclass;i++) { log_slx_capture(k,h,iyr,i) = -100;}
              if (it==2) for (int i=1;i<=nclass;i++) { log_slx_retaind(k,h,iyr,i) = -100; log_slx_discard(k,h,iyr,i) = log(1-TINY); }
            }
            break;
          case SELEX_PRESPECIFIED:
            for (int iyr=syr;iyr<=nyrRetro+nprojUse+1;iyr++){ #<1>
              int offset = (it-1)*nfleet*nsex*(nyr+1-syr+1)+nfleet*(nyr+1-syr+1)*(h-1)+(nyr+1-syr+1)*(k-1)+iyr-syr+1;
              if (it==1) for (int i=1;i<=nclass;i++) { log_slx_capture(k,h,iyr,i) = log(CustomSelex(offset,i));}
              if (it==2) for (int i=1;i<=nclass;i++) { log_slx_retaind(k,h,iyr,i) = log(CustomSelex(offset,i)); log_slx_discard(k,h,iyr,i) = log(1.0 - CustomSelex(offset,i) + TINY); }
            }
            break;
          default:
            if (slx_type_in(ipnt2,k) < 0){
              // Mirror case
              for (int iyr=syr;iyr<=nyrRetro+nprojUse+1;iyr++){ #<1>
                if (it==1) { log_slx_capture(k,h,iyr) = log_slx_capture(-slx_type_in(ipnt2,k),h,iyr);}
                if (it==2) { log_slx_retaind(k,h,iyr) = log_slx_retaind(-slx_type_in(ipnt2,k),h,iyr); 
                          log_slx_discard(k,h,iyr) = log_slx_discard(-slx_type_in(ipnt2,k),h,iyr); }
                }              
            } else {
              //non-mirror case
              int ipnt1 = (it-1)*nsex+h;
              int ipnt2 = (it-1)*nsex*nfleet+(h-1)*nfleet+k;
              if (h==1 || slx_bsex_in(it,k)==1){
                for (int iyr=syr;iyr<=nyrRetro+nprojUse+1;iyr++){ #<1>
                  int isel_vec = nsel_pnt+iYrsSelChanges(ipnt2,iyr);
                  selx = sel_store(isel_vec);
                  if (it==1) { log_slx_capture(k,h,iyr) = selx;}
                  if (it==2) { log_slx_retaind(k,h,iyr) = selx; log_slx_discard(k,h,iyr) = log(1.0 - exp(selx) + TINY); }
                }
                nsel_pnt %+=% nSelVaries(ipnt2);
              } else {
                for (int iyr=syr;iyr<=nyrRetro+nprojUse+1;iyr++){ #<1>
                  if (it==1) selx = log_slx_capture(k,1,iyr);
                  if (it==2) selx = log_slx_retaind(k,1,iyr);
                  if (it==1) log_slx_capture(k,h,iyr) = selx;
                  if (it==2) { log_slx_retaind(k,h,iyr) = selx; log_slx_discard(k,h,iyr) = log(1.0 - exp(selx) + TINY); }
                }
              } //--if  
            } //-if
        } //-- switch
      } //--k loop
    } //--h loop
  } //--it loop
```

1. NOTE: `iyr` runs `syr` to `nyrRetro`+`nprojUse`+1

Check for one fleet "inside" another: 

```{r, eval=FALSE}
  // Check for one fleet inside another 
  for (int it=1;it<=1;it++){
    for (int h=1;h<=nsex;h++){
      for (int k=1;k<=nfleet;k++){
        if (slx_include_in(1,k) > 0){
          for (int iyr=syr;iyr<=nyrRetro+nprojUse+1;iyr++) #<1>
            log_slx_capture(k,h,iyr) %+=% log_slx_capture(slx_include_in(1,k),h,iyr);
        }
      }
    }
  }
```

1. NOTE: `iyr` runs `syr` to `nyrRetro`+`nprojUse`+1

## in `calc_brute_equilibrium`

```{r, eval=FALSE}
  dvar_vector ssb_prj(syr,nyr+ninit);   ///> projected spawning biomass  #<1>
  dvar_vector hist_ssb(syr,nyrRetro);   ///> Historical SSB
```

 1. What's `ninit`?

## in `project_biomass`

```{r, eval=FALSE}
  dvar_vector ssb_prj(syr,nyr+iproj);  ///> projected spawning biomass #<1.
  dvar_vector hist_ssb(syr,nyrRetro);  ///> Historical SSB
```
1. What's `iproj`?

```{r, eval=FALSE}
  numbers_proj_gytl.initialize();
  for ( int ig = 1; ig <= n_grp; ig++ ) numbers_proj_gytl(ig)(1)(1) = d4_N(ig)(nyrRetro+1)(1);   #<1>
   
  // copy SSB into projection SSB
  hist_ssb = calc_ssb();                                                                         #<2>
  ssb_prj.initialize();
  for (int i = syr; i<=nyrRetro;i++) ssb_prj(i) = hist_ssb(i);                                   #<3>
  
  // recruitment distribution (constant recruitment)
  if (nsex>1){
    rtt(1) = Rbar(1) * rec_sdd(1);
    rtt(2) = Rbar(2) * rec_sdd(2);
  } else {
    rtt(1) = Rbar(1) * rec_sdd(1);
  }

  if (nsex>1){
    TotalSex1 = 0; TotalAll = 0;
    for (int i=YrRefSexR1;i<=YrRefSexR2;i++){
      TotalSex1 %+=% 1 / (1 + mfexp(-logit_rec_prop(i)));
      TotalAll %+=% 1;
    }
    SexRatio = TotalSex1/TotalAll;
  }
   
  // Now project forward (Rbar is adusted by the expected Rbar by year)
  for ( int i = 1; i <= iproj; i++ ){                                                          #<4>
    if (Stock_rec_prj==1) {                          // Constant recruitment
      for (int h=1; h<=nsex;h++) Rec_use(h) = fut_recruits(h,i)*exp(logRbar(nyrRetro+i));
    }
    if (Stock_rec_prj==RICKER){                      // Ricker
      ssb_use = ssb_prj(nyrRetro+i-Age_at_rec_prj);
      TotalRec = SR_alpha_prj*ssb_use*exp(-1*SR_beta_prj*ssb_use)*fut_recruits(1,i)*exp(logRbar(nyrRetro+i));
      if (nsex>1){
        Rec_use(1) = TotalRec * SexRatio;
        Rec_use(2) = TotalRec * (1 - SexRatio);
      } else {
        Rec_use(1) = TotalRec;
      }
    }
    if (Stock_rec_prj==BEVHOLT){                      // Beveton-Holt
      ssb_use = ssb_prj(nyr+i-Age_at_rec_prj);
      TotalRec = SR_alpha_prj*ssb_use/(SR_beta_prj+ssb_use)*fut_recruits(1,i)*exp(logRbar(nyrRetro+i));
      if (nsex>1){
        Rec_use(1) = TotalRec * SexRatio;
        Rec_use(2) = TotalRec * (1 - SexRatio);
      } else {
        Rec_use(1) = TotalRec;
      }
    }
    if (Stock_rec_prj==MEAN_RECRUIT){                    // Mean recruitment
      TotalRec = mean_rec_prj;
      if (nsex>1){
        Rec_use(1) = TotalRec * SexRatio;
        Rec_use(2) = TotalRec * (1 - SexRatio);
      } else {
        Rec_use(1) = TotalRec;
      }
    }
	 
    // Store start-year N-at-size
    d4_Pass.initialize();
    for ( int ig = 1; ig <= n_grp; ig++ )  d4_Pass(ig) = numbers_proj_gytl(ig,i,1); 

    // Compute the ABC and OFL
    FederalStuff = compute_OFL_and_ABC(nyrRetro+i, YrRefGrow, YrRefM1, YrRefM2, YrRefSea1, YrRefSea2, YrRefSel1, YrRefSel2, Rbar, d4_Pass);
    
    // Should the project consider the state TAC
    if (Apply_StateHCR_prj==1){     
      // Compute the State TAC
      StateTAC = CalcStateTAC(i,iproj,YrRefGrow);

      // Adjust so that the TAC is not larger than the ABC
      if (FederalStuff(3) < StateTAC)
       { TACType = 1; TargetC = FederalStuff(2); }                    /// > Hit the total catch OFL
      else
       { TACType = 2; TargetC = StateTAC; }                           /// > Hit the retained catc OFL
      mcoutDIAG << nyrRetro+i << " OFL " << FederalStuff(1) << " Total ABC " << FederalStuff(2) << " Retained ABC " << FederalStuff(3) << " StateTAC " << StateTAC << " Final Decision: " << TACType << " " << TargetC << endl;
    } else {
      TACType= 1; TargetC = FederalStuff(2);
      mcoutDIAG << nyrRetro+i << " OFL " << FederalStuff(1) << " Total ABC " << FederalStuff(2) << " Retained ABC " << FederalStuff(3) << " Final Decision: " << TACType << " " << TargetC << endl;
    }

    // DO a projection for the current F
    log_fimpbarPass = log_fimpbar;
    mcoutDIAG << nyrRetro+i << " Default F " << exp(log_fimpbarPass) << endl;
    project_one_year(i, iproj, YrRefGrow, YrRefM1, YrRefM2, YrRefSea1, YrRefSea2, YrRefSel1, YrRefSel2, rtt, rec_sdd, Rec_use, d4_Pass);
    mcoutDIAG << nyrRetro+i << " Default C (ret, total, by fleet) " << catch_pass << endl;
	
     int NeedToTune;
     if (Apply_HCR_prj==1) NeedToTune = YES;
     if (Apply_HCR_prj==0) NeedToTune = NO;
     if (TACType==1 && catch_pass(2)-1.0e-5 < TargetC) NeedToTune = NO;
     if (TACType==2 && catch_pass(1)-1.0e-5 < TargetC) NeedToTune = NO;

     ofstream manage_strat1("quota.inp");
     manage_strat1 << catch_pass << endl;
     manage_strat1 << "Here" << endl;
     manage_strat1 << catch_pass << endl;
     manage_strat1 << i << endl;
     manage_strat1.close();
     
     // Call the management startegy
     //int xx = system("dir");
     //cout << xx << endl;
     ifstream manage_strat2("quota.inp");
     float TAC_used;
     manage_strat2 >> TAC_used;
     cout << TAC_used << endl;
     manage_strat2.close();

     // Apply bisection to find the target F for the directed fishery
     if (NeedToTune==YES)
      {
       Fmin = 0; Fmax =1;
       for (int ii=1; ii<=20;ii++)
        {
         Fmult = (Fmin+Fmax)/2.0;
         for (int k=1;k<=nfleet;k++)
          if (Ffixed(k) != 1) log_fimpbarPass(k) = log(mfexp(log_fimpbar(k))*Fmult);
         project_one_year(i, iproj, YrRefGrow, YrRefM1, YrRefM2, YrRefSea1, YrRefSea2, YrRefSel1, YrRefSel2, rtt, rec_sdd, Rec_use, d4_Pass);
         if (TACType == 1)
          {
           if (catch_pass(2) > TargetC) Fmax = Fmult; else Fmin = Fmult;
          }
         if (TACType == 2)
          {
           if (catch_pass(1) > TargetC) Fmax = Fmult; else Fmin = Fmult;
          }
        }
       mcoutDIAG << nyrRetro+i << "CTUNE " << Fmult << " " << TargetC << " " << catch_pass << exp(log_fimpbarPass) << endl;
       mcoutDIAG << nyrRetro+i << "Final F " << exp(log_fimpbarPass) << endl;
       mcoutDIAG << nyrRetro+i << "Final C (ret, total, by fleet) " << catch_pass << endl;
      }
	  
     // Store catches during the projection
     for (int k=1;k<=nfleet+2;k++) catch_summary(k,i) = catch_pass(k);  #<5>
     catch_summary(nfleet+3,i) = FederalStuff(1);                       #<5>
     catch_summary(nfleet+4,i) = FederalStuff(2);                       #<5>
     catch_summary(nfleet+5,i) = StateTAC;                              #<5>
     ssb_prj(nyrRetro+i) = ssb_pass;                                    #<6>

    } // i = 1:iproj

  // return
  ssb_pass = ssb_prj(nyrRetro+iproj); #<7>
```

1. `numbers_proj_gytl` start at `nyrRetro`+1
2. `hist_ssb` runs syr:nyrRetro
3. filling up "historical portion" of prj_ssb
4. projection loop i: 1->iproj
5. `catch_summary`'
6. filling in rest of `ssb_proj` to nyrRetro + iproj
7. saving final projected ssb in nyrRetro+iproj as "global" variable `ssb_pass`

## doProjections_FinalPhase

  * dvar_vector `Bproj`(syr,nyr+nproj) ///> Biomass outout
  * loops for projected years run `i`=1:nproj with values assigned to $nyrRetro+iproj$

## write_eval

  * dvar_vector Bproj(syr,nyr+nproj)

```{r, eval=FALSE}
    mcoutPROJSSB << "BMSY ";
    for (int i=syr;i<=nyr+nproj;i++) mcoutPROJSSB << "SSB_" << i << " ";
    mcoutPROJSSB << endl;
    mcoutPROJCAT << "Draw Replicate F_val Fleet ";
    for (int i=syr;i<=nyr+nproj;i++) mcoutPROJCAT << "Catch_" << i << " ";
    mcoutPROJCAT << endl;

    #...

    for (int i=syr;i<=nyrRetro;i++) mcoutSSB << "SSB_" << i << " ";
    mcoutSSB << endl;
    for (int i=syr;i<=nyrRetro;i++) mcoutREC << "Rec_" << i << " ";
    mcoutREC << endl;

    #..

    for (int i=syr;i<=nyrRetro;i++) mcoutPROJSSB << ssb_old(i) << " ";
    for (int i=nyr+1;i<=nyrRetro+nproj;i++) mcoutPROJSSB <<  Bproj(i) << " ";
    #..
    if (k==3+nfleet || k==4+nfleet || k==5+nfleet) for (int i=syr;i<=nyr;i++) mcoutPROJCAT << "-1 ";
    if (k <=2+nfleet) mcoutPROJCAT << histcat(k);
    mcoutPROJCAT << catch_summary(k) << endl;
```

## calc_sdreport

```{r, eval=FALSE}
  dvar4_array ftmp(1,nsex,syr,nyrRetro,1,nseason,1,nclass);           ///> Fishing mortality
  dvar4_array ftmp2(1,nsex,syr,nyrRetro,1,nseason,1,nclass);          ///> Fishing mortality

      for (int h=1;h<=nsex;h++)
      for (int y=syr;y<=nyrRetro;y++)
        { ParsOut(Ipnt) = log(recruits(h,y)); Ipnt +=1; }

    for (int y=syr;y<=nyrRetro;y++)
      { ParsOut(NSSBPar+y-syr) = sd_log_ssb(y); }

   for ( int y = syr; y <= nyrRetro; y++ )  ParsOut(NfbarPar+y-syr) = mean(F(1,y));
```

```{r, eval=FALSE}
  // Zero catch
  F.initialize();
  for ( int h = 1; h <= nsex; h++ )
   for ( int i = syr; i <= nyrRetro; i++ )
    for ( int j = 1; j <= nseason; j++ )
     for ( int l = 1; l <= nclass; l++)
      F2(h,i,j,l) = 1.0e-10;
  calc_total_mortality();
  calc_initial_numbers_at_length();
  update_population_numbers_at_length();
  dyn_Bzero.initialize();
  dyn_Bzero = calc_ssb()(syr,nyrRetro);    #<1>

  if (OutDynB0==YES)
   for (int y=syr;y<=nyrRetro;y++)         #<1>
    ParsOut(NB0Par+y-syr) = log(dyn_Bzero(y)+0.00001); //--wts 20260918: added small constant to avoid log(0)
```

1. `dyn_Bzero` calculated for syr:nyrRetro here. Agrees w/ dimensioning but `NVarPar` expects it to be syr:nyr (see above)

## CreateOutput

```{r, eval=FALSE}
  dvar_matrix SummaryTable(syr,nyr,1,5);   #<1>
  dvar_matrix SummaryTableSD(syr,nyr,1,5); #<1>
```

1. Note time-dimesioned to syr:nyr

Likelihood_summary dataframe:

```{r, eval=FALSE}
  if (hasMMOD){
    for (int iy=syr; iy<=(nyrRetro+1);iy++){
    RepFile1 << "MMOD_data              " << "mmod_data         " << iy << "  NA      NA       NA      NA " <<
                                           nloglike(7,iy) << " " << mmodLkWgt             << " " << nloglike(7,iy)                             << endl;
    }
  }
```

InpOutFile5:

```{r, eval=FALSE}
  int IpntOut = 0;
  if (OutRecruit==YES) 
   {
    for (int h=1;h<=nsex;h++)
     for (int y=syr;y<=nyr;y++)
      {
       OutFile5 << h << " " << y << " " << "Log(rec)[" << h <<"," <<y<< "] " << ParsOut(NRecPar+IpntOut)<< " " << ParsOut.sd(NRecPar+IpntOut) << endl;
       SummaryTable(y,h) = ParsOut(NRecPar+IpntOut);
       SummaryTableSD(y,h) = ParsOut.sd(NRecPar+IpntOut);
       IpntOut += 1;
      }
     NparEst += IpntOut; 
   }  
  IpntOut = 0;   
  if (OutSSB==YES) 
   {
    for (int y=syr;y<=nyr;y++)
     {
      OutFile5 << y << " " << "Log(ssb)[" << y << "] " << ParsOut(NSSBPar+IpntOut)<< " " << ParsOut.sd(NSSBPar+IpntOut) << " " << endl;
      SummaryTable(y,3) = ParsOut(NSSBPar+IpntOut);
      SummaryTableSD(y,3) = ParsOut.sd(NSSBPar+IpntOut);
      IpntOut += 1;
     }
     NparEst += IpntOut; 
   }  
  IpntOut = 0;   
  if (Outfbar==YES)
   {
    for (int y=syr;y<=nyr;y++)
     {
      OutFile5 << 0 << " " << y << " " << "Mean(f)[" <<y<< "] " << ParsOut(NfbarPar+IpntOut)<< " " << ParsOut.sd(NfbarPar+IpntOut) << endl;
      SummaryTable(y,4) = ParsOut(NfbarPar+IpntOut);
      SummaryTableSD(y,4) = ParsOut.sd(NfbarPar+IpntOut);
      IpntOut += 1;
     }
     NparEst += IpntOut; 
   }  
  IpntOut = 0;   
  if (OutDynB0==YES)
   {
    for (int y=syr;y<=nyr;y++)
     {
      OutFile5 << 0 << " " << y << " " << "log(dyn ssb)[" <<y<< "] " << ParsOut(NB0Par+IpntOut)<< " " << ParsOut.sd(NB0Par+IpntOut) << endl;
      SummaryTable(y,5) = ParsOut(NB0Par+IpntOut);
      SummaryTableSD(y,5) = ParsOut.sd(NB0Par+IpntOut);
      IpntOut += 1;
     }
     NparEst += IpntOut; 
   }  
```

Summary dataframe:

```{r, eval=FALSE}
  for (int i=syr;i<=nyrRetro;i++)  #<1>
   {
    OutFile1 << i << " " << ssb(i) << " " << SummaryTable(i,3) << " " << SummaryTableSD(i,3) << " " << ssba(i) << " ";
    OutFile1 << dyn_Bzero(i) << " " << SummaryTable(i,5) << " " << SummaryTableSD(i,5) << " ";
    for (int h=1;h<=nsex;h++) OutFile1 << recruits(h,i) << " " << SummaryTable(i,h) << " " << SummaryTableSD(i,h) << " ";
    OutFile1 << SummaryTable(i,4) << " " << SummaryTableSD(i,4) << " ";
    for (int k=1;k<=2+nfleet;k++) OutFile1 << histcat(k,i) << " ";
    OutFile1 << rec_dev(i) << " " << logit_rec_prop(i) << " " << res_recruit(i);
    OutFile1 << endl;

    RepFile1 << i << " " 
             << ssb(i) << " " 
             << SummaryTable(i,3) << " " 
             << SummaryTableSD(i,3) << " " 
             << ssba(i) << " "
             << dyn_Bzero(i) << " " 
             << SummaryTable(i,5) << " " 
             << SummaryTableSD(i,5) << " ";
    for (int h=1;h<=nsex;h++) RepFile1 << recruits(h,i) << " " << SummaryTable(i,h) << " " << SummaryTableSD(i,h) << " ";
    RepFile1 << SummaryTable(i,4) << " " 
             << SummaryTableSD(i,4) << " ";
    for (int k=1;k<=2+nfleet;k++) RepFile1 << histcat(k,i) << " ";
    RepFile1 << rec_dev(i) << " " 
             << logit_rec_prop(i) << " " 
             << res_recruit(i)
             << endl;
   }
```

1. NOTE: note that this only outputs values of `SummaryTable` and `SummaryTableSD` for syr:nyrRetro although both are time-dimensioned syr:nyr

OutFile2 (Selectivity):

```{r, eval=FALSE}
  // Output the selex controls 
  OutFile2 << endl << "Selectivity" << endl; 
  for ( int h = 1; h <= nsex; h++ ) for ( int j = 1; j <= nfleet; j++ )
   OutFile2 << syr << " " << h << " " << j << " " << mfexp(log_slx_capture(j,h,syr)) << endl; #<1>
  for ( int h = 1; h <= nsex; h++ ) for ( int j = 1; j <= nfleet; j++ )
   OutFile2 << nyr << " " << h << " " << j << " " << mfexp(log_slx_capture(j,h,nyrRetro)) << endl;    #<2>
  OutFile2 << "retained" << endl; 
  OutFile2 << syr << " " << "1" << " " << "1" << " " << mfexp(log_slx_retaind(1,1,syr)) << endl;  
  OutFile2 << nyrRetro << " " << "1" << " " << "1" << " " << mfexp(log_slx_retaind(1,1,nyrRetro)) << endl;               
```

1. `log_slx_capture` time-dimensioned syr:nyrRetro(?)
2. year = `nyr` here is WRONG

OutFile2:

```{r, eval=FALSE}
  OutFile1 << "Year" << " "; for (int i=1;i<=nseason;i++) OutFile1 << i << " "; OutFile1 << endl;
  for (int iyr=syr;iyr<=nyrRetro;iyr++)
   { OutFile1 << iyr << " " << m_prop(iyr) << endl; }
```

OutFile1/Repfile1:

```{r, eval=FALSE}
  OutFile1 << endl << "Natural_mortality-by-class: dataframe" << endl;
  OutFile1 << "Year Sex Maturity "; for (int i=1;i<=nclass;i++) OutFile1 << i << " "; OutFile1 << endl;
  RepFile1 << endl << "#--------------------------------------------------------------------------------------------" << endl;
  RepFile1 << endl << "Natural_mortality-by-class: dataframe" << endl;
  RepFile1 << "year sex maturity " << strMidpoints << endl;
  for (int h=1;h<=nsex;h++)
   for (int m=1;m<=nmature;m++)
    for (int iyr=syr;iyr<=nyrRetro;iyr++) {                                                     #<1>
      OutFile1 << iyr << " " << sexes(h) << " " << maturestate(m) << " " << M(h,m,iyr) << endl; #<1>
      RepFile1 << iyr << " " << sexes(h) << " " << maturestate(m) << " " << M(h,m,iyr) << endl; #<1>
    }
  OutFile1 << ">EOD<" << endl;
  RepFile1 << ">EOD<" << endl << endl;

  OutFile1 << endl << "# Fully-selected fishing mortality-by-season by sex and fishery" << endl;//--misnomer: ft is capture rate, not fishing mortality rate
  OutFile1 << "Fully-selected_FM_by_season_sex_and_fishery: dataframe" << endl;
  OutFile1 << "Sex Fleet Year "; for (int i=1;i<=nseason;i++) OutFile1 << i << " "; OutFile1 << endl;
  RepFile1 << endl << "#--------------------------------------------------------------------------------------------" << endl;
  RepFile1 << "Fully-selected_capture_rate_by_season_sex_and_fishery: dataframe" << endl;
  RepFile1 << "sex fleet year "; for (int i=1;i<=nseason;i++) RepFile1 << i << " "; RepFile1 << endl;
  for (int h=1;h<=nsex;h++)
   for (int k=1;k<=nfleet;k++)
    for (int iyr=syr;iyr<=nyrRetro;iyr++) {  #<1>
      OutFile1 << sexes(h) << " " << fleetname(k) << " " << iyr << " " << ft(k,h,iyr) << endl; 
      RepFile1 << sexes(h) << " " << fleetname(k) << " " << iyr << " " << ft(k,h,iyr) << endl; 
    } 
  OutFile1 << ">EOD<" << endl;
  RepFile1 << ">EOD<" << endl << endl;

  OutFile1 << endl << "# Fully-selected_fishing mortality by fleet" << endl;//--misnomer: ft is capture rate, not fishing mortality rate
  OutFile1 << "Fully-selected_FM_by_fleet: dataframe" << endl;
  OutFile1 << "Sex Year Season "; for (int i=1;i<=nfleet;i++) OutFile1 << fleetname(i) << " "; OutFile1 << endl;
  for (int h=1;h<=nsex;h++)
   for (int i=syr;i<=nyrRetro;i++) #<1>
    for (int j=1;j<=nseason;j++)
     {
      OutFile1 << sexes(h) << " " << i << " " << j << " ";
       for (int k=1;k<=nfleet;k++) OutFile1 << ft(k,h,i,j) << " ";
      OutFile1 << endl;
     }
  OutFile1 << ">EOD<" << endl;

  RepFile1 << endl << "#--------------------------------------------------------------------------------------------" << endl;
  calc_predicted_capture_abdbio();
  RepFile1 << "Predicted_capture_abundance-at-size: dataframe" << endl;//--fishery captured abundance by year, fleet, sex, and size
  RepFile1 << "year fleet sex " << strMidpoints << endl;
  for (int i=syr;i<=nyrRetro;i++){ #<1>
    for (int f=1;f<=nfleet;f++) { 
      for (int h=1;h<=nsex;h++)
        RepFile1 << i << " " << fleetname(f) << " " << sexes(h) << " " <<  pre_capabd(i,f,h) << endl; 
    }
  } 
  RepFile1 << ">EOD<" << endl << endl;

  RepFile1 << endl << "#--------------------------------------------------------------------------------------------" << endl;
  RepFile1 << "Predicted_capture_biomass-at-size: dataframe" << endl;//--fishery captured biomass by year, fleet, sex, and size
  RepFile1 << "year fleet sex " << strMidpoints << endl;
  for (int i=syr;i<=nyrRetro;i++){ #<1>
    for (int f=1;f<=nfleet;f++) { 
      for (int h=1;h<=nsex;h++)
        RepFile1 << i << " " << fleetname(f) << " " << sexes(h) << " " <<  pre_capbio(i,f,h) << endl; 
    }
  } 
  RepFile1 << ">EOD<" << endl << endl;

  OutFile1 << endl << "# Fishing mortality-at-size by sex and season (Continuous)" << endl;//--correctly includes retention and discard mortality components
  OutFile1 << "FM-at-size_by_sex_and_season_(continuous): dataframe" << endl;
  OutFile1 << "Sex Year Season " << strMidpoints << endl;
  RepFile1 << endl << "#--------------------------------------------------------------------------------------------" << endl;
  RepFile1 << "FM-at-size_by_sex_and_season_(continuous): dataframe" << endl;//--correctly includes retention and discard mortality components
  RepFile1 << "sex year season " << strMidpoints << endl;
  for (int h=1;h<=nsex;h++)
   for (int iyr=syr;iyr<=nyrRetro;iyr++) #<1>
    for (int is=1;is<=nseason;is++) { 
      OutFile1 << sexes(h) << " " << iyr << " " << is << " " << F(h,iyr,is) << endl; 
      RepFile1 << sexes(h) << " " << iyr << " " << is << " " << F(h,iyr,is) << endl; 
    } 
  OutFile1 << ">EOD<" << endl;
  RepFile1 << ">EOD<" << endl << endl;

  OutFile1 << endl << "# Fishing mortality-at-size by sex and season (Discrete)" << endl;
  OutFile1 << "FM-at-size_by_sex_and_season_(discrete): dataframe" << endl;
  OutFile1 << "Sex Year Season "; for (int i=1;i<=nclass;i++) OutFile1 << i << " "; OutFile1 << endl;
  for (int h=1;h<=nsex;h++)
   for (int iyr=syr;iyr<=nyrRetro;iyr++) #<1>
    for (int is=1;is<=nseason;is++)
     { OutFile1 << sexes(h) << " " << iyr << " " << is << " " << F2(h,iyr,is) << endl; } 
  OutFile1 << ">EOD<" << endl;

  
  OutFile1 << endl << "# Total mortality by size-class (continuous)" << endl;
  OutFile1 << "TM-by-size-class_(continuous): dataframe" << endl;
  OutFile1 << "Sex Maturity Year Season Total_mortality " << strMidpoints << endl;
  RepFile1 << endl << "#--------------------------------------------------------------------------------------------" << endl;
  RepFile1 << "TM-by-size-class_(continuous): dataframe" << endl;
  RepFile1 << "sex maturity year season " << strMidpoints << endl;
  for (int h=1;h<=nsex;h++)
   for(int m=1;m<=nmature;m++)
    for (int i=syr;i<=nyrRetro;i++) #<1>
     for (int j=1;j<=nseason;j++){
      OutFile1 << sexes(h) << " " << maturestate(m) << " " << i << " " << j << " " << Z(h,m,i,j) << endl;
      RepFile1 << sexes(h) << " " << maturestate(m) << " " << i << " " << j << " " << Z(h,m,i,j) << endl;
     }
  OutFile1 << ">EOD<" << endl;
  RepFile1 << ">EOD<" << endl;
  
  OutFile1 << endl << "# Total mortality by size-class (discrete)" << endl;
  OutFile1 << "TM-by-size-class_(discrete): dataframe" << endl;
  OutFile1 << "Sex Maturity Year Season Total_mortality "; for (int i=1;i<=nclass;i++) OutFile1 << i << " "; OutFile1 << endl;
  for (int h=1;h<=nsex;h++)
   for( int m=1;m<=nmature;m++)
    for (int i=syr;i<=nyrRetro;i++) #<1>
     for (int j=1;j<=nseason;j++)
      OutFile1 << sexes(h) << " " << m << " " << i << " " << j << " " << Z2(h,m,i,j) << endl;
  OutFile1 << ">EOD<" << endl;
```

1. writing out for syr:nyrRetro

### Total numbers-at-size

All dimensioned syr:(nyrRetro+1) to include start of "next" year:

```{r, eval=FALSE}
  // Print total numbers at length
  dvar_matrix N_initial(1,n_grp,1,nclass);
  dvar_matrix N_total(syr,nyrRetro+1,1,nclass);
  dvar_matrix N_males(syr,nyrRetro+1,1,nclass);
  dvar_matrix N_females(syr,nyrRetro+1,1,nclass);
  dvar_matrix N_males_new(syr,nyrRetro+1,1,nclass);
  dvar_matrix N_females_new(syr,nyrRetro+1,1,nclass);
  dvar_matrix N_males_old(syr,nyrRetro+1,1,nclass);
  dvar_matrix N_females_old(syr,nyrRetro+1,1,nclass);
  dvar_matrix N_males_mature(syr,nyrRetro+1,1,nclass);
  dvar_matrix N_females_mature(syr,nyrRetro+1,1,nclass);
  dvar_matrix N_males_imature(syr,nyrRetro+1,1,nclass);
  dvar_matrix N_females_imature(syr,nyrRetro+1,1,nclass);
  dvar_matrix N_males_imature_new(syr,nyrRetro+1,1,nclass);
  dvar_matrix N_males_imature_old(syr,nyrRetro+1,1,nclass);
  dvar_matrix N_males_mature_new(syr,nyrRetro+1,1,nclass);
  dvar_matrix N_males_mature_old(syr,nyrRetro+1,1,nclass);
  dvar_matrix N_females_imature_new(syr,nyrRetro+1,1,nclass);
  dvar_matrix N_females_imature_old(syr,nyrRetro+1,1,nclass);
  dvar_matrix N_females_mature_new(syr,nyrRetro+1,1,nclass);
  dvar_matrix N_females_mature_old(syr,nyrRetro+1,1,nclass);
```

### OutInpFile1 (output ctl file)

```{r, eval=FALSE}
  OutInpFile1<< "# pre-specified selectivity/retention (ordered by type, sex, fleet and year)" << endl;
  for (int it=1;it<=2;it++){
    for (int h=1;h<=nsex;h++){
      for (int k=1;k<=nfleet;k++){
        if (slx_type_in((it-1)*nsex+h,k) == SELEX_PRESPECIFIED){
          OutInpFile1 << "# Pre-specified values for " << seltypes(it) << " for " << sexes(h) <<"s for " <<fleetname(k) << endl;
          for (int ii=1;ii<=nclass;ii++) OutInpFile1 << "# " << setw(9) << setfixed() << ii << " "; OutInpFile1 << endl;
          int offset = (it-1)*nfleet*nsex*(nyr+1-syr+1)+nfleet*(nyr+1-syr+1)*(h-1)+(nyr+1-syr+1)*(k-1);
          for (int iyr =syr;iyr<=nyr+1;iyr++){
            for (int ii=1;ii<=nclass;ii++) OutInpFile1 << setw(9) << setprecision(7) << setfixed() << CustomSelex(offset+iyr-syr+1,ii) << " ";
            OutInpFile1 << "# " << iyr << endl;
          }
        }// if 
      }// k loop
    } // h loop
  }// it loop
```

```{r CodeNotes-NumberOfParams, child=file.path("./qq00a-GMACS-CodeNotes.NumerberOfParameters.qmd"),eval=TRUE}
```
